\Title

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
Methods

Here, we keep the analogy with spiking neurons, and we try to observe the detection of spiking motifs using the spatiotemporal kernels. If so, this precise spatiotemporal patterns prove to be of interest for neural computations and one can infer that biological neurons make use of this information as well. The 3D convolution represents the linear integration of the spike train as the linear input to the neuron. The activation function of our model is then a softmax function that implements a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, analogous to a spiking winner-takes-all network~\citep{nessler_bayesian_2013}. In our MLR model, a probability value for each class (i.e., each direction of motion) is predicted for each position $x, y$ and time $\timev$ as a softmax function of the linear combination of the list of events. Formally, using the kernels, the input raster plot is transformed into a probability with the following formula:
%
\begin{linenomath*}
\begin{equation}\label{eq:mlr}
    \forall x, y, \timev, \forall \class \in [1, N_\class], \; \;
Pr(k=\class \; \vert \;  x, y, \timev) =
\frac {\exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class) }{Z(x, y, \timev)}
\end{equation} 
\end{linenomath*}
%
where $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp (\mathcal{C}_\class(x, y, \timev) +\bias_\class)$ is the partition function, and $\bias_\class$ is the bias associated with the class $\class$. 
In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to high postsynaptic activity, making it progressively more likely to generate an output spike. The spiking output of the model corresponds to an event with the highest probability class.

Now that this general framework has been explained, we can add some heuristics, based on neuroscientific observations, to constrain our model and its strategies for solving the task used to solve the ecological task described in section~\ref{sec:task}. Note that the general framework is similar to that presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve and a deeper analysis of the results, these novel neuroscience-inspired heuristics are the main methodological differences with that previous study. First, in our simulations, we set the size of the kernels to $(\Ktime, \Kx, \Ky) = (21, 17, 17)$ and define as many classes as the number of motions (directions and velocities): $N_\class = XXX8 \times XXX3$. To avoid introducing biases in the directions, we apply a circular mask to the spatial dimensions of the kernels. We also included a prior in the selectable motions, as there is a prior for slow speeds in natural scenes~\citep{vacher_bayesian_2018}. Since we want to capture the possible convergence of the trajectories of the events converging on each voxel, we apply a mask to the spatiotemporal kernels such that the smaller the delay, the smaller the radius of the circular mask that is applied (\seeFig{kernels} for an illustration). This is consistent with neuroscientific principles because, due to the limits of conduction delay along horizontal connections, the synaptic delay is physically related to the distance between the presynaptic and postsynaptic neurons. Second, we observed that moving images produced trajectories of ON and OFF spikes, and that these were present in both polarities. This is due to the fact that our whitened images have a symmetry in the luminance profiles, and an image with inverted contrast is indistinguishable from a standard one. Since this arrangement of polarities is independent of motion, we added a mechanism that collects the linear values for the movie and the movie with the ON and OFF cells flipped, keeping only the maximum value for each voxel. This is similar to the computation done for complex cells in primary visual cortex.

>>>>

Here, we keep the analogy with spiking neurons, and we try to observe the detection of spiking motifs using the spatiotemporal kernels. If so, this precise spatiotemporal patterns prove to be of interest for neural computations and one can infer that biological neurons make use of this information as well. The 3D convolution represents the linear integration of the spike train as the linear input to the neuron. The activation function of our model is then a softmax function that implements a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, analogous to a spiking winner-takes-all network~\citep{nessler_bayesian_2013}. In our MLR model, a probability value for each class (i.e., each direction of motion) is predicted for each position $x, y$ and time $\timev$ as a softmax function of the linear combination of the list of events. Formally, using the kernels, the input raster plot is transformed into a probability with the following formula:
%
\begin{linenomath*}
\begin{equation}\label{eq:mlr}
    \forall x, y, \timev, \forall \class \in [1, N_\class], \; \;
Pr(k=\class \; \vert \;  x, y, \timev) =
\frac {\exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class) }{Z(x, y, \timev)}
\end{equation} 
\end{linenomath*}
%
where $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp (\mathcal{C}_\class(x, y, \timev) +\bias_\class)$ is the partition function, and $\bias_\class$ is the bias associated with the class $\class$. 
In particular, we expect that some specific motifs may generate tightly synchronized outputs as they are integrated in the basal dendritic tree, leading to high postsynaptic activity, making it progressively more likely to generate an output spike. The spiking output of the model corresponds to an event with the highest probability class.

Now that this general framework has been explained, we can include some heuristics based on neuroscientific observations to constrain our model and its strategies for solving the ecological task described in section~\ref{sec:task}. Note that the general framework is similar to that presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve and a deeper analysis of the results, these novel neuroscience-inspired heuristics are the main methodological differences with that previous study. First, in our simulations, we set the size of the kernels to $(\Ktime, \Kx, \Ky) = (21, 17, 17)$ and define as many classes as the number of motions (directions and velocities): $N_\class = XXX8 \times XXX3$. To avoid introducing biases in the directions, we apply a circular mask to the spatial dimensions of the kernels. We also included a prior in the selectable motions, as there is a prior for slow speeds in natural scenes~\citep{vacher_bayesian_2018}. Since we want to capture the possible convergence of the trajectories of the events converging on each voxel, we apply a mask to the spatiotemporal kernels such that the smaller the delay, the smaller the radius of the circular mask that is applied (\seeFig{kernels} for an illustration). This is consistent with neuroscientific principles because, due to the limits of conduction delay along horizontal connections, the synaptic delay is physically related to the distance between the presynaptic and postsynaptic neurons. Second, we observed that moving images produced trajectories of ON and OFF spikes, and that these were present in both polarities. This is due to the fact that our whitened images have a symmetry in the luminance profiles, and an image with inverted contrast is indistinguishable from a standard one. Since this arrangement of polarities is independent of motion, we added a mechanism that collects the linear values for the movie and the movie with the ON and OFF cells flipped, keeping only the maximum value for each voxel. This is similar to the computation done for complex cells in primary visual cortex.
